Apparent preferences for cognitive effort fade when multiple forms of effort

### 2020-11-03 Cost_manuscript.Rmd
slight methods section restructuring

---
title: Apparent preferences for cognitive effort fade when multiple forms of effort
  and delay are interleaved in a foraging environment
author: "Claudio Toro-Serey, Gary Kane, & Joseph T. McGuire"
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{float}
- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}} }
output:
  html_document:
    df_print: paged
  pdf_document:
    highlight: haddock
    keep_tex: no
  word_document: default
csl: apa.csl
---

@@

  Evaluating whether to pursue a desired outcome often requires assessing the demands required for obtaining it. In psychology and economics, demands for both time and effort have traditionally been identified as costs that detract from the net subjective value of rewards (Westbrook & Braver, 2015). Indeed, decision makers tend to prefer low-effort and immediate rewards over larger rewards that demand longer delays and higher effort (Ainslie, 1975; Hull, 1943; Kool & Botvinick, 2018; Kool, McGuire, Rosen, & Botvinick, 2010; Treadway, Buckholtz, Schwartzman, Lambert, & Zald, 2009; Westbrook, Kester, & Braver, 2013). However, people often change their attitudes towards exerting effort in everyday decisions, resulting in pursual of effortful avenues instead. Emerging experimental and theoretical work has chiefly explained this phenomenon by suggesting a reciprocal relationship between effort and outcome, whereby exerting effort boosts the perceived value of its outcomes (Kacelnik and Marsh, 2002; Mochon, Norton, and Ariely, 2012; Hernandez Lallement et al., 2014), and positive outcomes imbue value into their preceding effortful behavior (i.e. learned industriousness, Eisenberger, 1992). Though intuitively valid and empirically supported, this explanation leaves open questions about what intrinsic features of an effortful task can make it attractive outside in its own right, and how contextual factors mold its perceived costs (Inzlicht, Shenhav, and Olivola, 2018). Understanding these qualities could improve our ability to motivate completion of effortful daily activities, such as schoolwork or physical exercise.

@@

  In this study, we tested the hypothesis that preferences would vary depending on whether individuals face a single or multiple types of behavioral demands (i.e. cognitive effort, physical effort, and delay). We leveraged the ecological and computational richness of foraging paradigms to provide new insights on a number of open questions (Inzlicht, Shenhav, and Olivola, 2018). First, the existence of a well-defined optimal choice strategy in foraging designs helped us identify when demands boost the value of rewards (i.e. overharvesting poor offers). Second, increasing the level of exposure to types of demands across two experiments allowed us to asses whether their cost was intrinsic or relative. Finally, confronting participants with both mental and physical effort let us probe whether these preferences were domain general or task-specific, and in the case of the latter, isolate factors that influence preferences beyond the often-cited associative relationship between effort and outcome.

  We found mixed support for our pre-registered hypotheses, which are detailed in https://osf.io/2rsgm/. Confirming our predictions, we found that individuals could approximate optimality by preferring higher rewards and shorter delays, regardless of the richness of the environment or whether they faced a single or intermixed demands. When faced with a single type of demand, people in the cognitive effort group accepted more trials than optimal, whereas physical effort and wait groups under-accepted, partially matching our predictions and highlighting that cognitive effort can add value to outcomes. However, against our expectations, we found that a similar preference for cognitive effort disappeared over time as decision makers foraged for types of demands in addition to rewards. Computational modeling suggested that cognitive effort was valuable because it shortened the subjective elapsed time of trials, a perception that could be recalibrated as individuals experienced multiple forms of demands. The shortened time sensitivity of cognitive effort could be due to its constant recruitment of working memory, impeding estimations of the nominal time, or because of the discrete nature of the task (versus a continuous configuration of the physical and wait tasks). These results expand our understanding of the attractiveness of effort, bringing forth outcome-independent features that can be leveraged to motivate effort engagement in diverse everyday scenarios.

@@

  All experimental tasks were implemented using PsychoPy 2 (v1.85.1, CITE) on a Macbook Pro laptop. In this task, participants foraged for monetary rewards in an environment in which each trial required either physical effort or cognitive effort for a set period of time (the “handling time”), or an equivalent unfilled delay (Figure 1). Their goal was to maximize their gains within a fixed amount of time. On each trial a monetary offer was displayed for 2 s, and participants had the opportunity to expend time and/or effort during the handling time in order to earn it. Upon completion of a trial participants saw a 2 s window displaying the reward obtained, which was followed by a travel time to the next offer. Alternatively, the participant could quit at any point during the handling time by pressing the spacebar on the computer, and immediately start traveling.

  Each participant was assigned to one of three types of costs (cognitive effort, physical effort, or delay), and a fourth group of equal size faced an effortless physical task that involved minimal gripping. Each group was unaware that other cost conditions existed. Participants exerted physical effort by maintaining grip on a handheld dynamometer (XXXX, BIOPAC Systems, United States) using their dominant hand. Gripping requirements were calibrated at 20% of maximum voluntary contraction (MVC, acquired at the beginning of the session). Cognitive effort entailed switching among Stroop, dot motion coherence, and flanker tasks. In Stroop, one of three color names was displayed on the screen (red, blue or green), with a font color that was either congruent (e.g. word red painted in red) or incongruent (e.g. word red painted in blue). Participants had to select the color of the font, not the word displayed (i.e. they had to suppress their tendency to read the word). For motion coherence, 100 solid white dots moved on the screen. A fraction of these dots moved cohesively to the left or right, while the rest moved in random directions (coherence could be either 30% or 40%, uniformly sampled). Participants had to respond with the direction of the cohesive set of dots. In flanker, rows of arrowheads pointed either to the left or the right (maximum of 3 rows, 3 to 13 arrowheads per row). Participants responded with the direction of the center arrowhead, which could point in the same or opposite direction from its neighbors. These tasks were configured so that responses always involved a left or right key press (e.g. for Stroop, two colored circles were presented at each side of the screen). During the handling time, cognitive tasks and their configurations were randomly sampled, and were presented for 1 s followed by a 1 s inter-stimulus interval. Participants were asked to respond within each task’s presentation time. Before the experiment, participants trained in each cognitive task until they correctly performed six consecutive tasks of each kind. If participants failed to maintain above-threshold gripping or made two mistakes during the cognitive task, they were forced to travel and missed the reward.

  There were three block types, in which handling times of 2, 10, or 14 seconds were paired with travel times of 14, 6, and 2 seconds, respectively (note that all combinations add up to 16 seconds). Timing parameters were held constant within each 7-min block. Each of these blocks was experienced in a pseudo-randomized order, and repeated in the same order after a short break in order to probe choice stability (total session length = 42 minutes). Reward amounts varied uniformly per trial (4, 8, or 20 cents), with the constraint that every reward was presented twice every six trials. This prevented sequences from being dominated by a single amount during any window of time. Timing information was disclosed at the beginning of each block, and rewards displayed during a 2 s offer window before each trial began. Participants received training prior to the experimental session, and were told about all possible environmental statistics in order to preclude experience-dependent learning.

@@

  Foraging theory posits that accepting a delayed reward should depend on the opportunity cost (OC) of time incurred in obtaining it, given by the richness of the environment (Stephens and Krebs, 1977). In this study, the richness of the environment was manipulated by the length of the handling and travel times, where shorter travel times produced richer environments. Since time combinations were fixed per block, we calculated each block’s optimal accept/reject strategy by computing all decision strategies according to the following equation:

$\dfrac {\sum p_i R_i} {\sum p_i H_i + 3T},p \in \{0, 1\}$

  where *R*, *H*, and *p* are the reward, handling time, and acceptance probability of offer i, respectively, and T is the travel time. This gave us the total reward per second attainable in each block as a function of the acceptance threshold (i.e. the lowest amount accepted). Figure 2 shows possible earnings for each choice strategy, as well as the lowest amount participants should accept in order to maximize their rewards (circled dots). For example, a participant in a 10 s handling block should accept 8-cent and 20-cent rewards (and reject 4-cent rewards) to maximize their reward rate. Note that accepting every offer was often detrimental to participant earnings. Significant deviations from the optimal strategy can be understood as reflecting changes in subjective OC. This approach allowed us to test if certain demands boosted the value of offers, expressed as a subjective OC that fell below the optimal rate.

@@

  All analyses were performed in R 3.5.2 (CITE). First we tested whether decision makers integrated delay and reward information. To address the prediction that participants would be more likely to accept higher reward and shorter handling time trials, we fit a mixed-effects logistic regression to predict trial-wise acceptances, using a random intercept to account for biases in participant behavior. We included regressors for handling time and reward, as well as two covariates that probed the influence of recent history on choices. The first was an autoregressive regressor containing the number of consecutive offer misses prior to a given trial (misses could be due to quits or forced travels). The second regressor tracked the sum of the previous *n* offers. We identified the *n* that minimized Akaike's Information Criterion (AIC) among 6 versions of the model (tracing from 2 previous trials up to 7, beyond which the model failed to converge), and reported all coefficients from the winning model.

  Our foraging task was configured such that over and under accepting were detrimental to total earnings. To confirm this, we fit a general linear model with constant, linear, and quadratic terms to estimate the correspondence between proportion of trials completed (independent variable) and total earnings (dependent variable). A significant quadratic coefficient thus would signal that the task statistics operated as expected. Next, to determine the optimality of each group’s decisions, we performed two-sided one-sample t-tests to see if the proportion of acceptances for each time/reward combination was significantly different from the optimal rate (see Operationalization of Cocitst). This resulted in 36 independent tests (3 rewards amounts, 3 handling/travel time combinations, and 4 groups), so we corrected for multiple comparisons using False Discovery Rate (FDR, CITE Y&..).

@@

  In order to further look at the effect of handling, offer, and cost type, we computed the probability of accepting a trial with a mixed-effects logistic regression, using the lme4 package (CITE). Based on the task structure and our main question, our a priori model of interest included cost condition, handling time, and reward amount as fixed main effects, and participant ID as a random intercept. Cost condition was modeled with three categorical terms, with the fourth condition as the reference condition. We ran three versions of the model with different reference conditions, in order to test all pairwise differences among the four cost conditions. (**exponentiate coefficients to make them more interpretable?**) We then examined whether this a priori model outperformed both simpler and more complex models. We used both AIC and Bayesian Information Criterion (BIC) to determine the model that minimized the negative log-likelihood while penalizing the addition of parameters. The regression with each combination of predictors was fitted in a forward order: 1) intercept only; 2) condition only; 3) handling time only; 4) offer amount only; 5) handling time and offer amount (added to the preregistered comparison); 6) condition, handling time, and reward main effects (a priori model from above); 7) adding a handling-by-reward interaction; and 8) all possible two-way interactions. We predicted that model 6 (the a priori setup) would have the lowest AIC and BIC. Nested models with similar AIC were statistically compared using an analysis-of-deviance. The significance test was computed as the probability of the reduction in deviance, based on a chi-square distribution with degrees of freedom equal to the difference in the number of parameters between models.

  Before modeling each individual's overarching subjective OC, we examined whether participants were consistent and stable in their choices. For consistency, we visually examined survival curves indicating at what point during the handling time participants quit a trial (censored points included trial completions and forced travels), in addition to quitting time distributions. Stability was assessed by computing each participant’s total proportion of acceptances pre- and post-midpoint for every block, comparing the mean proportion of acceptances using paired permutations (5000 iterations). We then tested whether the relative differences in costs observed using data from the whole experimental session were still present on each half of the experiment. The winning model from the mixed-effects logistic comparison was applied to each half separately. In each case, the reference cost category was rotated in order to assess pairwise differences among all costs. We evaluated the consistency of the relative costs by visually comparing the patterns of significant differences between experimental halves.

  Finally, we fitted the subjective OC of each type of demand (see Operationalization of Cost). OC was computed as the product of a free parameter (gamma) and the handling time. Gamma can be interpreted as the subjective value of each second spent in an environment. If a trial's reward rate (i.e. offer amount divided by the handling time) fell below gamma, a participant was better off finding more profitable alternatives. In other words, individuals with higher gammas required higher rewards in order to accept a given offer. We used a softmax function to model each participant’s probability of completing a trial based on the difference between the delayed reward’s magnitude and the estimated OC for each cost type

$P(accept)_i = \dfrac {1} {1 + e^{-\beta(R_i - \gamma H_i)}}$

  Both gamma ($\gamma$) and the inverse temperature parameter ($\beta$) of the softmax function were estimated independently for each participant, using a maximum likelihood (MLE) approach through the NLOPTR package (CITE **or grid search?**). The gamma optimization search space was bound by extreme choice values. For example, someone with an extremely high OC would not accept even the most beneficial offers, and would reject a 20 cent offer when the handling time was 2 seconds (or 20/2). Conversely, a participant experiencing low opportunity costs would accept the non-beneficial offer of 4 cents for a 14 second delay (or 4/14). We cross-validated each subject’s OC value using the pre-midpoint data for estimation, and post-midpoint choices for testing. The estimates from the training sample were used to predict acceptances in the testing sample, and the mean percent correctly predicted was reported for each group. We then used a two-sided one-sample t-tests to compare each group's gammas to the average optimal earning rate (0.74, see Operationalization of Cost), allowing us to determine whether any demand boosted an offer's value. Finally, we determined which cost type produced the highest discounting by comparing gammas with an ANOVA (with condition as a factor), followed by post-hoc permutations (5000 iterations) comparing mean gammas per group.

@@

  The original between-subjects task was modified so that participants foraged for cost types in addition to rewards. Every participant experienced all forms of cost (delay, physical effort, and cognitive effort). There were two block types, in which delay trials were interspersed with either physical-effort or cognitive-effort trials. This setup prevented participants from having to rapidly switch between response modalities across trials (i.e. keyboard press and handgrip). Block combinations (i.e. physical/wait and cognitive/wait) were experienced three times by each participant in 7-minute-long, interleaved blocks, thus matching the experiment’s length to the between-subject version. Half of the participants experienced a block sequence that started with the physical block. Timing parameters were matched to the middle condition from Experiment 1 (i.e. 10 s handling and 6 s travel), allowing a balance between acquiring enough observations per trial type and being able to compare results across experiments. Participants were informed that this timing combination would be constant prior to beginning the experiment. Cost type combinations were disclosed at the beginning of each block, and reward/cost offers displayed before each trial begins (e.g. “8 cents for physical effort”). Unlike the previous experiment, participants could express their decision to quit during the offer window (instead of during the handling time). Participants trained in each type of demand until they reached the same criteria as in the between-subject experiment.

@@

  To compare acceptance rates among cost conditions, we first computed the probability of accepting a trial with a mixed-effects logistic regression. The model included cost condition and reward amount as fixed main effects, and participant ID as a random intercept (handling time is not modeled, as participants experienced a single time combination). Cost condition was modeled with three categorical terms, with the fourth condition as the reference condition (two effort conditions, and two delay conditions corresponding to each effort type). We ran three versions of the model with different reference conditions, in order to test all relevant pairwise differences among the four cost conditions. We compared the a priori model to alternative parameterizations using AIC and BIC, varying model complexity in a forward order: 1) intercept only; 2) cost only; 3) reward only; 4) cost and reward main effects (a priori model); and 5) adding a two-way interaction. We predicted that model 4 would have the last considerable decrease in the negative log-likelihood. Models with similar AIC were formally compared using the analysis-of-deviance approach from the between-subjects experiment.

  Next we probed the confidence and consistency with which participants made choices. First, we qualitatively assessed the number of quits performed during during the handling time through a survival curve, and plotted the quit response time distribution as a function of block to examine choice confidence over time. Next, we computed the proportion of quits that were performed during the choice window versus during the handling time. We interpreted as confidence if a participant made at least 80% of their choices during the offer window. We then computed each participant’s total proportion of acceptances on the first two and last two blocks, and compared them using a paired permutation analysis (5000 iterations). We avoided the third and fourth block because their effort designation was counterbalanced across participants, which would have biased our estimates.

  **old** To estimate the subjective opportunity cost of each condition, we used a logistic function to model each participant’s probability of accepting a trial based on the difference between the delayed reward’s magnitude and the estimated opportunity cost (OC) for each cost type. As before, OC was computed as the product of a free parameter (gamma) and the handling time. Both gamma and the temperature parameter of the logistic function were estimated at the subject level, independently for each subject and cost type. We cross-validated each subject’s OC value using the first four blocks (two cognitive, two physical) for estimation, and the last two blocks (one of each effort type) choices for testing. The estimates were used to predict acceptances in the testing sample, and the mean percent correctly predicted was reported for each cost type. The OC estimates for each group were compared using a repeated measures ANOVA with cost as a factor. We performed pairwise paired permutations as post-hoc tests.

**new** The adaptive nature of choices in this experiment demanded more flexibility than the single-parameter model fitted above. Based on the results across experiments, an adequate model of behavior should meet X criteria: 1) show deviations from optimality that were not simply due to overharvesting 2) persistently higher acceptances for cognitive effort when faced with single behavioral costs; 3) slight reduction of acceptances in effortful conditions (**fatigue?**); 4) an initial preference for cognitive effort trials when foraging for costs and rewards 5) this preference disappears as acceptances decreased to a common rate across costs; 6) participants kept track of recent offer history; 7) effects were robust to block order effects. Alternative foraging models from the literature were not fitted, as they did not qualitatively meet these criteria (wikenheiser, C&D, Fawcett...)

@@

  These findings provide a first glimpe at differences in cost, showing that individuals in the wait condition accepted the least and earned the most. Moreover, they suggest that costs associated with cognitive effort can boost the value of an offer instead, even when accepting it negatively impacts total earnings.

@@

  Understanding what drives the sporadic attractiveness of effortful options remains elusive. In this study, we tested the hypothesis that such appeal could arise from being able to compare different means to achieve the same outcome. Using two foraging experiments that exposed individuals to a single or intermixed types of demands, we found 1) that preferences for mental effort are stable when individuals face a single form of demand, but fade over time when other costs are also experienced; and 2) the phenomenon can be explained by a model in which preferences depend on how participants construe the amount of time a task takes, which could hypothetically be recalibrated when all tasks are interleaved.

>>>>

---
title: Apparent preferences for cognitive effort fade when multiple forms of effort
  and delay are interleaved in a foraging environment
author: "Claudio Toro-Serey, Gary Kane, & Joseph T. McGuire"
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{float}
- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}} }
output:
  pdf_document:
    highlight: haddock
    keep_tex: no
  word_document: #default
    reference_docx: Style_reference_draft.docx
  html_document:
    df_print: paged
csl: apa.csl
---

@@

  Evaluating whether to pursue a desired outcome often requires assessing the demands required for obtaining it. In psychology and economics, demands for both time and effort have traditionally been identified as costs that detract from the net subjective value of rewards (Westbrook & Braver, 2015). Indeed, decision makers tend to prefer low-effort and immediate rewards over larger rewards that demand longer delays and higher effort (Ainslie, 1975; Hull, 1943; Kool & Botvinick, 2018; Kool, McGuire, Rosen, & Botvinick, 2010; Treadway, Buckholtz, Schwartzman, Lambert, & Zald, 2009; Westbrook, Kester, & Braver, 2013). However, people often change their attitudes towards exerting effort in everyday decisions, resulting in pursual of effortful avenues instead. Emerging experimental and theoretical work has chiefly explained this phenomenon by suggesting a reciprocal relationship between effort and outcome, whereby exerting effort boosts the perceived value of its outcomes (Kacelnik and Marsh, 2002; Mochon, Norton, and Ariely, 2012; Hernandez Lallement et al., 2014), and positive outcomes imbue value into their preceding effortful behavior (i.e. learned industriousness, Eisenberger, 1992). Though intuitively valid and empirically supported, this explanation leaves open questions about what features of an effortful task can make it attractive in its own right, and how contextual factors mold its perceived costs (Inzlicht, Shenhav, and Olivola, 2018). Understanding these intrinsic qualities could improve our ability to motivate completion of effortful daily activities, such as schoolwork or physical exercise.

@@

  In this study, we tested the hypothesis that preferences would vary depending on whether individuals face a single or multiple types of behavioral demands (i.e. cognitive effort, physical effort, and delay). We explored this question in two foraging experiments: a between-subjects design in which participants foraged for rewards while performing a single demand, and a within-subjects extension that had individuals forage for demands in addition to rewards. We leveraged the ecological and computational richness of foraging paradigms to provide new insights on a number of open questions (Inzlicht, Shenhav, and Olivola, 2018). First, the existence of a well-defined optimal choice strategy helped us identify when demands boost the value of rewards (i.e. overharvesting poor offers). Second, increasing the level of exposure to types of demands allowed us to asses whether their cost was intrinsic or relative. Finally, confronting participants with both mental and physical effort let us probe whether these preferences were domain general or task-specific, and in the case of the latter, isolate factors that influenced preferences beyond the often-cited associative relationship between effort and outcome.

  Our results provided mixed support for our pre-registered hypotheses, which are detailed in https://osf.io/2rsgm/. Confirming our predictions, we found that individuals could approximate optimality by preferring higher rewards and shorter delays, regardless of the richness of the environment or whether they faced a single or intermixed demands. When faced with a single type of demand, people in the cognitive effort group accepted more trials than optimal, whereas physical effort and wait groups under-accepted, partially matching our predictions and highlighting that cognitive effort can add value to outcomes. However, against our expectations, we found that a similar preference for cognitive effort disappeared over time as decision makers foraged for types of demands in addition to rewards. Computational modeling suggested that cognitive effort was valuable because it shortened the subjective elapsed time of trials, a perception that could be recalibrated as individuals experienced multiple forms of demands. The shorter time sensitivity of cognitive effort could be due to its constant recruitment of working memory, impeding estimations of the nominal time, or because of the discrete nature of the task (versus a continuous configuration of the physical and wait tasks). These results expand our understanding of the attractiveness of effort, bringing forth outcome-independent features that can be leveraged to motivate effort engagement in diverse everyday scenarios.

@@

  All experimental tasks were implemented using PsychoPy 2 (v1.85.1, Pierce2019) on a Macbook Pro laptop. In this task, participants foraged for monetary rewards in an environment in which each trial required either physical effort or cognitive effort for a set period of time (the “handling time”), or an equivalent unfilled delay (Figure 1). Their goal was to maximize their gains within a fixed amount of time. On each trial a monetary offer was displayed for 2 s, and participants had the opportunity to expend time and/or effort during the handling time in order to earn it. Upon completion of a trial participants saw a 2 s window displaying the reward obtained, which was followed by a travel time to the next offer. Alternatively, the participant could quit at any point during the handling time by pressing the spacebar on the computer, and immediately start traveling.

  Each participant was assigned to one of three types of costs (cognitive effort, physical effort, or delay), and a fourth group of equal size faced an effortless physical task that involved minimal gripping. Each group was unaware that other cost conditions existed. Participants exerted physical effort by maintaining grip on a handheld dynamometer (XXXX, BIOPAC Systems, United States) using their dominant hand. Gripping requirements were calibrated at 20% of maximum voluntary contraction (MVC, acquired at the beginning of the session). Cognitive effort entailed switching among Stroop, dot motion coherence, and flanker tasks. In Stroop, one of three color names was displayed on the screen (red, blue or green), with a font color that was either congruent (e.g. word red painted in red) or incongruent (e.g. word red painted in blue). Participants had to select the color of the font, not the word displayed (i.e. they had to suppress their tendency to read the word). For motion coherence, 100 solid white dots moved on the screen. A fraction of these dots moved cohesively to the left or right, while the rest moved in random directions (coherence could be either 30% or 40%, uniformly sampled). Participants had to respond with the direction of the cohesive set of dots. In flanker, rows of arrowheads pointed either to the left or the right (maximum of 3 rows, 3 to 13 arrowheads per row). Participants responded with the direction of the center arrowhead, which could point in the same or opposite direction from its neighbors. These tasks were configured so that responses always involved a left or right key press (e.g. for Stroop, two colored circles were presented at each side of the screen). During the handling time, cognitive tasks and their configurations were randomly sampled, and were presented for 1 s followed by a 1 s inter-stimulus interval. Participants were asked to respond within each task’s presentation time. Before the experiment, participants trained in each cognitive task until they correctly performed six consecutive tasks of each kind. While in the handling time, if participants failed to maintain above-threshold gripping or made two mistakes during the cognitive task, they were forced to travel and missed the reward.

  There were three block types, in which handling times of 2, 10, or 14 seconds were paired with travel times of 14, 6, and 2 seconds, respectively (note that all combinations add up to 16 seconds). Timing parameters were held constant within each 7-min block. Each of these blocks was experienced in a pseudo-randomized order, and repeated in the same order after a short break in order to probe choice stability (total session length = 42 minutes). Reward amounts varied uniformly per trial (4, 8, or 20 cents), with the constraint that every reward was presented twice every six trials. This prevented sequences from being dominated by a single amount during any window of time. Timing information was disclosed at the beginning of each block, and rewards displayed during a 2 s offer window before each trial began. Participants received training prior to the experimental session, and were told about all possible environmental statistics in order to preclude biases due to experience-dependent learning (Dundon2020; Garrett2020).

@@

  Foraging theory posits that accepting a delayed reward should depend on the opportunity cost (OC) of time incurred in obtaining it, given by the richness of the environment (Stephens2019). In this study, the richness of the environment was manipulated by the length of the handling and travel times, where shorter travel times produced richer environments. Since time combinations were fixed per block, we calculated each block’s optimal accept/reject strategy by computing all decision strategies according to the following equation:

$\dfrac {\sum p_i R_i} {\sum p_i H + 3T},p \in \{0, 1\}$

  where *R* and *p* are the reward, and acceptance probability of offer i, respectively, *H* is the handling time for the block, and *T* its associated travel time. This gave us the total reward per second attainable in each block as a function of the acceptance threshold (i.e. the lowest amount accepted). Figure 2 shows possible earnings for each choice strategy, as well as the lowest amount participants should accept in order to maximize their rewards (circled dots). For example, a participant in a 10 s handling block should accept 8-cent and 20-cent rewards (and reject 4-cent rewards) to maximize their reward rate. Note that accepting every offer was often detrimental to participant earnings. Significant deviations from the optimal strategy can be understood as reflecting changes in subjective OC. This approach allowed us to test if certain demands boosted the value of offers, expressed as a subjective OC that fell below the optimal rate.

@@

  All analyses were performed in R 3.5.2 (RTeam2018). First we tested whether decision makers integrated delay and reward information. To address the prediction that participants would be more likely to accept higher reward and shorter handling time trials, we fit a mixed-effects logistic regression to predict trial-wise acceptances (using the lme4 package, Bates2015), giving a random intercept to each subject to account for biases in participant behavior. We included regressors for handling time and reward, as well as two covariates that probed the influence of recent history on choices. The first was an autoregressive regressor containing the number of consecutive offer misses prior to a given trial (misses could be due to quits or forced travels). The second regressor tracked the sum of the previous *n* offers. We identified the *n* that minimized Akaike's Information Criterion (AIC) among 6 versions of the model (tracing from 2 previous trials up to 7, beyond which the model failed to converge), and reported all coefficients from the winning model.

  Our foraging task was configured such that over and under accepting were detrimental to total earnings. To confirm this, we fit a general linear model with constant, linear, and quadratic terms to estimate the correspondence between proportion of trials completed (independent variable) and total earnings (dependent variable). A significant quadratic coefficient thus would signal that the task statistics operated as expected. Next, to determine the optimality of each group’s decisions, we performed two-sided one-sample t-tests to see if the proportion of acceptances for each time/reward combination was significantly different from the optimal rate (see Operationalization of Cost). This resulted in 36 independent tests (3 rewards amounts, 3 handling/travel time combinations, and 4 groups), so we corrected for multiple comparisons using False Discovery Rate (FDR, Benjamini2001).

@@

  In order to further look at the effect of handling, offer, and cost type, we computed the probability of accepting a trial with a mixed-effects logistic regression. Based on the task structure and our main question, our a priori model of interest included cost condition, handling time, and reward amount as fixed main effects, and a random intercept per subject. Cost condition was modeled with three categorical terms, with the fourth condition as the reference condition. We ran three versions of the model with different reference conditions, in order to test all pairwise differences among the four cost conditions. We then examined whether this a priori model outperformed both simpler and more complex models. We used both AIC and Bayesian Information Criterion (BIC) to determine the model that minimized the negative log-likelihood while penalizing the addition of parameters. The regression with each combination of predictors was fitted in a forward order: 1) intercept only; 2) condition only; 3) handling time only; 4) offer amount only; 5) handling time and offer amount (added to the preregistered comparison); 6) condition, handling time, and reward main effects (a priori model from above); 7) adding a handling-by-reward interaction; and 8) all possible two-way interactions. We predicted that model 6 (the a priori setup) would have the lowest AIC and BIC. Nested models with similar AIC were statistically compared using an analysis-of-deviance. The significance test was computed as the probability of the reduction in deviance, based on a chi-square distribution with degrees of freedom equal to the difference in the number of parameters between models.

  Next, we examined whether participants were consistent and stable in their choices. For consistency, we visually examined survival curves indicating at what point during the handling time participants quit a trial (censored points included trial completions and forced travels), in addition to quitting time distributions. We interrogated stability by computing each participant’s total proportion of acceptances pre- and post-midpoint for every block type, comparing the mean proportion of acceptances using paired permutations (5000 iterations). We then tested whether the observed differences in costs were still present on each half of the experiment by applying winning model from the mixed-effects logistic comparison to each half separately. In each case, the reference cost category was rotated in order to assess pairwise differences among all costs. We evaluated the consistency of the relative costs by visually comparing the patterns of significant differences between experimental halves.

@@

  The original between-subjects task was modified so that participants foraged for cost types in addition to rewards. Every participant experienced all forms of cost (delay, physical effort, and cognitive effort). There were two block types, in which delay trials were interspersed with either physical-effort or cognitive-effort trials. This setup prevented participants from having to rapidly switch between response modalities across trials (i.e. keyboard press and handgrip). Block combinations (i.e. physical/wait and cognitive/wait) were experienced three times by each participant in 7-minute-long, interleaved blocks, thus matching the experiment’s length to the between-subject version. Half of the participants experienced a block sequence that started with the physical block. Timing parameters were matched to the middle condition from Experiment 1 (i.e. 10 s handling and 6 s travel), establishing a balance between acquiring enough observations per trial type and being able to compare results across experiments. Participants were informed that this timing combination would be constant prior to beginning the experiment. Cost type combinations were disclosed at the beginning of each block, and reward/cost offers displayed before each trial begins (e.g. “8 cents for physical effort”). Unlike the previous experiment, participants could express their decision to quit during the offer window (instead of during the handling time). Participants trained in each type of demand until they reached the same criteria as in the between-subject experiment.

@@

  To compare acceptances among cost conditions, we first computed the probability of accepting a trial with a mixed-effects logistic regression. The model included cost condition and reward amount as fixed main effects, and participant ID as a random intercept (handling time is not modeled, as participants experienced a single time combination). Cost condition was modeled with three categorical terms, with the fourth condition as the reference condition (two effort conditions, and two delay conditions corresponding to each effort type). We ran three versions of the model with different reference conditions, in order to test all relevant pairwise differences among the four cost conditions. We compared the a priori model to alternative parameterizations using AIC and BIC, varying model complexity in a forward order: 1) intercept only; 2) cost only; 3) reward only; 4) cost and reward main effects (a priori model); and 5) adding a two-way interaction. We predicted that model 4 would have the last considerable decrease in the negative log-likelihood. Models with similar AIC were formally compared using the analysis-of-deviance approach from the between-subjects experiment.

  Next we probed consistency and stability in people's choices. First, we qualitatively assessed the number of quits performed during during the handling time through a survival curve, and plotted the quit response time distribution as a function of block to examine choice consistency over time. Next, we computed the proportion of quits that were performed during the choice window versus during the handling time. We interpreted as confidence if a participant made at least 80% of their choices during the offer window. We then computed each participant’s total proportion of acceptances on the first two and last two blocks, and compared them using a paired permutation analysis (5000 iterations). We avoided the third and fourth block because their effort designation was counterbalanced across participants, which would have biased our estimates.

### **Computational Modeling of Foraging Behavior**

  We begun the modeling approach by fitting a fixed subjective OC for each type of demand (see Operationalization of Cost), which acted as a base model. OC was computed as the product of a free parameter (gamma) and the handling time. Gamma can be interpreted as the subjective value of each second spent in an environment. If a trial's reward rate (i.e. offer amount divided by the handling time) fell below gamma, a participant was better off finding more profitable alternatives. In other words, individuals with higher gammas required higher rewards in order to accept a given offer. We used a softmax function to model each participant’s probability of completing a trial based on the difference between the delayed reward’s magnitude and the estimated OC for each cost type

$P(accept)_i = \dfrac {1} {1 + e^{-1/\beta(R_i - \gamma H_i)}}$

  Both gamma ($\gamma$) and the temperature parameter ($\beta$) of the softmax function were estimated independently for each participant, using a maximum likelihood (MLE) approach through the NLOPTR package (CITE **GARY INFO**). The gamma optimization search space was bound by extreme choice values. For example, someone with an extremely high OC would not accept even the most beneficial offers, and would reject a 20 cent offer when the handling time was 2 seconds (or 20/2). Conversely, a participant experiencing low opportunity costs would accept the non-beneficial offer of 4 cents for a 14 second delay (or 4/14). We cross-validated each subject’s OC value using the pre-midpoint data for estimation, and post-midpoint choices for testing. The estimates from the training sample were used to predict acceptances in the testing sample, and the mean percent correctly predicted was reported for each group. We then used a two-sided one-sample t-tests to compare each group's gammas to the average optimal earning rate (**0.74**, see Operationalization of Cost), allowing us to determine whether any demand boosted an offer's value. Finally, we determined which cost type produced the highest discounting by comparing gammas with an ANOVA (with condition as a factor), followed by post-hoc permutations (5000 iterations) comparing mean gammas per group.

  The adaptive nature of choices in this experiment demanded more flexibility than the single-parameter model fitted above. Based on the results across experiments, an adequate model of behavior should meet X criteria: 1) show deviations from optimality that were not simply due to overharvesting 2) persistently higher acceptances for cognitive effort when faced with single behavioral costs; 3) slight reduction of acceptances in effortful conditions (**fatigue?**); 4) an initial preference for cognitive effort trials when foraging for costs and rewards 5) this preference disappears as acceptances decreased to a common rate across costs; 6) participants kept track of recent offer history; 7) effects were robust to block order effects. Alternative foraging models from the literature were not fitted, as they did not qualitatively meet these criteria (wikenheiser, C&D, Fawcett...)

@@

  These findings provide a first glimpe at differences in cost, showing that individuals in the wait condition accepted the least and earned the most. Moreover, they suggest that costs associated with cognitive effort can boost the value of an offer instead, as reflected by higher acceptances at the expense of earnings.

@@

  Understanding the sporadic attractiveness of effortful options remains elusive. In this study, we tested the hypothesis that such appeal could arise from being able to compare different means to achieve the same outcome. Using two foraging experiments that exposed individuals to a single or intermixed types of demands, we found that 1) preferences for cognitive effort are stable when individuals face a single form of demand, but fade over time when other costs are also experienced; and 2) the phenomenon can be explained by a model in which preferences depend on how participants construe the amount of time a task takes, which could hypothetically be recalibrated when all tasks are interleaved.
